---
title: "animated histo_waffle"
author: "A Gruer"
date: "20 November 2018"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(gganimate)
library(lubridate)
library(magrittr)

theme_set(theme_minimal())
```



```{r get data}
wait_data <- read_csv("wait_data.csv", col_types = "DicDDc")
```
Monthly snapshots of patients waiting for a procedure.

Data dictionary

  - date_for: month of snapshot
  - days_on_list: number of days since the patient was added to the list
  - month_event: one of four events to describe what happened to the patient during the month
  - added date: date patient was added to the list
  - removed_date: date patient was removed from the list, NA if patient not yet removed
  - pat_id: unique identifier for wach patient
   
  

Month events
In any one month a patient can have one of four events

 1. added to the list 
 1. added_and_removed in the same month
 1. removed from the list
 1. waited for the whole month

```{r table state_by_month}
table(wait_data$date_for, wait_data$month_event)
```



For a single month create a static "histo_waffle" and compare to a histogram

```{r static histo_waffle}

wait_data %>%
  filter(date_for == as_date("2018/01/31")) %>%

  # bin the waiting days into bins 20 days wide
  # group by the bin and within each bin asssign a row number to each individual
  mutate(bin  = cut_width(days_on_list, 20, center = 10)) %>%
  group_by(bin) %>%
  mutate(row_id = row_number(days_on_list)) %>%
  ungroup() %>%

  # some regex to get just the right hand limit of each bin as a number
  # and use that as the leabel for the bins
  # number following a comma.
  # comma is a non-matching group
  # number is group 1
  # str_match returns a matrix - column one is the entire match.
  # column 2 is group 1 and so on
  mutate(bin = fct_relabel(bin, ~ str_match(.x, "(?:,)(\\d+)")[, 2])) %>%

  ggplot(aes(bin, row_id)) +
  geom_tile(width = 0.9, height = 0.9) +
  coord_equal()  +
  scale_x_discrete(drop = FALSE, name = "days waited") +
  scale_y_continuous(breaks = seq.int(2, 10, 2), name = "number of patients") +
  theme(panel.grid =  element_blank()) #+
  #annotate("text", x =10, y = 6, label ="waiting list")

```    

```{r static histogram}     
  
wait_data %>%
  filter(date_for == as_date("2018/01/31") ) %>%
  ggplot(aes(days_on_list)) +
  geom_histogram(binwidth = 20, center = 10)
```


Animation!!

```{r animate, fig.path="histo_waffle.gif"}
wait_data %>%

  # filter out added_and_removed in same month - too distracting
  # filter out removed - for now the pat_id will just disappear in the
  # month the patient is removed
  filter(month_event %in% c("waited", "added")) %>% 
  
  # bin the waiting days into bins 20 days wide
  # group by the date_for bin and within each bin asssign a row number to each
  # individual
  mutate(bin  = cut_width(days_on_list, 20, center = 10)) %>% 
  group_by(date_for, bin) %>% 
  mutate(row_id = row_number(days_on_list)) %>% 
  ungroup() %>% 
    
  # some regex to get just the right hand limit of each bin as a number
  # and use that as the leabel for the bins
  # number following a comma. 
  # comma is a non-matching group
  # number is group 1
  # str_match returns a matrix - column one is the entire match.
  # column 2 is group 1 and so on
  mutate(bin = fct_relabel(bin, ~ str_match(.x, "(?:,)(\\d+)")[, 2])) %>%

  #plotting code - TODO automate manual breaks
  ggplot(aes(bin, row_id, group = pat_id)) +
  geom_tile( width = 0.9, height = 0.9, fill = "mediumblue", alpha = 0.5) + 
  scale_x_discrete(breaks = seq.int(20, 720, 40),
                   drop = FALSE,
                   name = "days waited") +
  scale_y_continuous(breaks = seq.int(2, 20, 2), name = "number of patients") +
  coord_equal() +
  theme(panel.grid =  element_blank())   +
  labs(x = "days waited (up to)", title = "{format(frame_time, '%B %y')}")  +

  ## ANIMATE
  # frames
  transition_components(id = pat_id,
                          #time = month  ,
                          time =  as_date(date_for),
                          enter_length = as_date(20),
                          exit_length =  as_date(20)  
                          ) +
  
  # entries and exits
  enter_grow(fade = TRUE) + 
  exit_shrink(fade = TRUE) +
  
  # easing of points algorithm
  ease_aes("cubic-in-out")  
    

 
 #save to file
  #gganimate::anim_save("wl.gif") 
  

  #transition_manual(frames = date_for) +
  #labs(x = "days waited (up to)", title = "{current_frame}")
  
   # 
   # transition_states(
   # date_for,
   # transition_length = 1,#0.5,
   # state_length = 0.5,#0.0001,
   # wrap = FALSE) +
   # labs(x = "days waited (up to)", title = "{closest_state}")
  
  # transition_time(
  #   as_date( date_for) )+
   # labs(x = "days waited (up to)", title = "{frame_time}")

  # transition_events(start = as_date(date_for)) +
  # labs(x = "days waited (up to)", title = "{frame_time}")
```



